\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation

### 2017-10-27 OaklandSubmission/intro.tex
nit
intro

Since its introduction in 1980s, there has been a long sequence
of work that has transformed \mpc from a mere theoretical concept to
something practically efficient and usable. To understand the
state-of-the-art, let us consider the following example underlying many
secure prediction algorithms. Suppose Alice wants to write
a \mpc protocol to securely
compute $w^Tx >b$. Here $w$ (a vector) and $b$ (a scalar) constitute
the server model, and $x$ is the client's input vector. Further,
$\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
inner product of  $w^{T}$ and $x$. With the current state-of-the-art,
Alice has following options:

@@

This paper presents \tool\footnote{Read as ``easy peasy'',
stands for Easy 2 Party Computation}, a new \mpc framework.
\tool is the first
``cryptographic-cost aware'' framework that generates efficient and
scalable \mpc protocols from high-level programs devoid of any
cryptographic details. The generated protocols comprise combinations
of arithmetic and boolean circuits and have performance comparable to
or better than the custom, specialized
protocols from previous work. Below we describe the
salient features of \tool.

\subsubsection*{Ease of programming} \tool source programs are ideal
functionalities that describe ``what'' computation needs to be done,
rather than ``how'' to do it. In particular, the programmer writes the
high-level computation without thinking about the underlying
cryptographic details. For example, an \tool source
program for $w^Tx >b$ is as follows:

\vspace{0.2cm}
\begin{lstlisting}[language=C]
uint w[30] = input1(); uint b = input1();
uint x[30] = input2();
uint acc = 0;
for i in [0:30] { acc =  acc + (w[i] * x[i]); }
output((acc > b) ? 1 : 0)
\end{lstlisting}

\vspace{0.2cm}
The program is quite similar to what a programmer might
write in C++ or Java. The simplicity of the
language comes with its usual benefits: it is easily accessible to the
developers, there are fewer avenues for making mistakes, developers
don't bear the burden of getting cryptographic details right, and it
is easy to maintain and modify the programs. Needless to say,
frameworks that expose low-level circuit APIs to the programmer do not
enjoy these benefits.

\subsubsection*{Cryptogaphic-cost aware compiler} The \tool compiler
compiles a source program to a hybrid computation consisting
of \emph{public} and \emph{secret} parts. In the example above, for
instance, \tool compiler realizes that the array index~\ls{i} is
public, and hence, the array accesses need not be compiled as
multiplexers. Further, within the secret parts, \tool compiler is
aware
of the cryptographic costs of arithmetic and boolean representations
of the source language operators. Based on these costs, the compiler
automatically picks arithmetic or boolean
representations for different sub-parts, and generates the
corresponding circuits along with the required inter-conversion
gates. The outcome is an effcient \mpc protocol combining arithmetic
and boolean circuits, while the programmer remains
oblivious of all these cryptographic details. Our evaluation
shows that the performance of these automatically generated protocols
is comparable
to or better than the custom, specialized protocols developed in the
previous work. Indeed, \tool is the first such cryptographic-cost
aware compiler.

@@

We have implemented the \tool toolchain using ABY~\cite{aby} as the
cryptographic back-end. We evaluate \tool by implementing a wide range
of secure prediction benchmarks including linear and na\"{i}ve bayes
classifiers, decision trees, deep neural networks, state-of-the-art
classifiers such as Tensorflow~\cite{tensorflow}
and \bonsai~\cite{bonsai}, and also the matrix factorization example
from Nikolaenko et al.~\cite{valeriaMatrix}. Our results demonstrate
three key points. First, \tool makes it convenient for general
programmers to write \mpc protocols. E.g. we provide the first \mpc
implementation of \bonsai~\cite{bonsai}, and it was programmed at a
high-level in \tool by a
cryptography novice. Second, the performance of the protocols
generated by \tool are comparable to or better than (up to 19x in some
cases) their state-of-the-art specialized implementations. Finally, we
show that with partitioning, we are able to implement applications
that require more than 100 million gates.

The rest of the paper is structured as follows. We formalize \tool and
prove its correctnss and security theorems in
Section~\ref{sec:ld}. Section~\ref{sec:pipe} details our partitioning
scheme and formally proves its security. Section~\ref{sec:impl}
and~\ref{sec:eval} present details of \tool implementation and
evaluation. Finally, Section~\ref{sec:related} talks about the related
work. We start with an overview of \tool.

@@

%% \noindent\textbf{Other Related Work.} 
%% Before ABY \cite{aby}, to improve efficiency, several works proposed
%% to mix secure computation protocols based on homomorphic
%% encryption and Yao's garbled circuits protocol, e.g., \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}
%%   and developed tools that allowed writing a mixed \mpc protocol, e.g., \cite{bogdanov,lone,tasty}. The breaking up of the function into homomorphic encryption and garbled circuits can
%% even be done automatically as was shown in \cite{autoS}. 
%% However, as was noted by ABY, due to high conversion cost between homomorphic encryption and Yao's garbled circuit, these mixed protocols do not gain much performance over a single protocol. Also, none of these works give a high-level developer-friendly language support.
%% No prior work has given a formal type system and operational semantics for mixed arithmetic and boolean back-ends.
%% We provide a detailed survey of related work in \sectionref{related}.

>>>>

Our goal is to build a framework that enables programmers, who are not
expert in cryptography, to easily program efficient and scalable \mpc
protocols. Since its introduction in 1980s, there has been a long
sequence of work along the same lines. To understand the
state-of-the-art, let us consider an example underlying many
secure prediction algorithms. Say Alice wants to write
a \mpc protocol to securely
compute $w^Tx >b$. Here $w$ (a vector) and $b$ (a scalar) constitute
the server model, and $x$ is the client's input vector. Further,
$\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
inner product of  $w^{T}$ and $x$. Alice has following options.

@@

This paper presents \tool\footnote{Read as ``easy peasy'',
stands for Easy 2 Party Computation.}, a new \mpc framework.
\tool is the first
``cryptographic-cost aware'' framework that generates efficient and
scalable \mpc protocols from high-level programs devoid of any
cryptographic details. The generated protocols comprise combinations
of arithmetic and boolean circuits and have performance comparable to
or better than the custom, specialized
protocols from previous work. Below we describe the
salient features of \tool.

\begin{figure}
\begin{lstlisting}[language=C]
uint w[30] = input1(); uint b = input1();
uint x[30] = input2();
uint acc = 0;
for i in [0:30] { acc =  acc + (w[i] * x[i]); }
output((acc > b) ? 1 : 0)
\end{lstlisting}
\caption{\tool code for $w^Tx >b$}
\label{fig:introexample}
\end{figure}

\subsubsection*{Ease of programming} \tool source programs are ideal
functionalities that describe ``what'' computation needs to be done,
rather than ``how'' to do it. In particular, the programmer writes the
high-level computation without thinking about the underlying
cryptographic details. For example, Figure~\ref{fig:introexample}
shows an \tool source
program for $w^Tx >b$.
%
%\vspace{0.2cm}
%
%\vspace{0.2cm}
The program is quite similar to what a programmer might
write in C++ or Java. The simplicity of the
language comes with its usual benefits: it is easily accessible to the
developers, there are fewer avenues for making mistakes, developers
don't bear the burden of getting cryptographic details right, and it
is easy to maintain and modify the programs. Needless to say,
frameworks that expose low-level circuit APIs to the programmer do not
enjoy these benefits.

\subsubsection*{Cryptogaphic-cost aware compiler} The \tool compiler
compiles a source program to a hybrid computation consisting
of \emph{public} and \emph{secret} parts. In the example above, for
instance, \tool compiler realizes that the array index~\ls{i} is
public, and hence, the array accesses need not be compiled
obliviously. Further, within the secret parts, \tool compiler is
aware
of the cryptographic costs of arithmetic and boolean representations
of the source language operators. Based on these costs, the compiler
automatically picks arithmetic or boolean
representations for different sub-parts, and generates the
corresponding circuits along with the required inter-conversion
gates. The outcome is an effcient \mpc protocol combining arithmetic
and boolean circuits, while the programmer remains
oblivious of all these cryptographic details. Our evaluation
shows that the performance of these automatically generated protocols
is comparable
to or better than the custom, specialized protocols developed in the
previous work. Indeed, \tool is the first such cryptographic-cost
aware compiler.

@@

\subsubsection*{Evaluation} We have implemented the \tool toolchain
using ABY~\cite{aby} as the
cryptographic back-end. We evaluate \tool by implementing a wide range
of secure prediction benchmarks including linear and na\"{i}ve bayes
classifiers, decision trees, deep neural networks, state-of-the-art
classifiers from Tensorflow~\cite{tensorflow}
and \bonsai~\cite{bonsai}, and also the matrix factorization example
from Nikolaenko et al.~\cite{valeriaMatrix}. Our results demonstrate
three key points. First, \tool makes it convenient for general
programmers to write \mpc protocols. E.g. we provide the first \mpc
implementation of \bonsai~\cite{bonsai}, and it was programmed in the
high-level \tool source language by a non-cryptographer. Second, the
performance
of the protocols
generated by \tool are comparable to or better than (up to 19x in some
cases) their state-of-the-art specialized implementations. Finally, we
show that with partitioning, we are able to implement applications
that require more than 300 million gates.

\subsubsection*{Related Work}
Before ABY~\cite{aby}, several works have proposed
combining secure computation protocols based on homomorphic
encryption and Yao's garbled circuits
(e.g. \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}),
  and some have also developed tools that allow writing such combinations
  (e.g. \cite{bogdanov,lone,tasty}). Kerschbaum et al.~\cite{autoS}
  even provide a scheme to automatically select optimal protocols
  for sub-compuations. However, as Demmler et al.~\cite{aby} note, due
  to high conversion cost between
homomorphic encryption and Yao's garbled circuits, these combined
protocols do not gain much performance over a single
protocol. Additionally, these prior works neither have a
developer-friendly language, nor do they formalize their type system
and operational semantics, and show them to be consistent.
We provide a detailed survey of related work in \sectionref{related}.

The rest of the paper is structured as follows. We formalize \tool and
prove its correctnss and security theorems in
Section~\ref{sec:ld}. Section~\ref{sec:pipe} details our partitioning
scheme and proves its security. Section~\ref{sec:impl}
presents the implementation details, and
Section~\ref{sec:eval} present a comprehensive evaluation of \tool. We
begin with an overview.

@@
